PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-13 paper/sec-results.tex
General revision

In this section we demonstrate the capabilities of \textsc{MG5\_aMC}+\textsc{PineAPPL} by
computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW,
for a representative set of processes in which EW corrections may have a
sizeable effect on the accuracy of the theoretical predictions.
In order to consider some realistic kinematics for these
processes, we rely on a set of measurements that are commonly included in PDF
fits. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We describe
the settings employed for the computations, and the corresponding results for
each process considered.

@@

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, photon-induced (PI) contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation. Monte Carlo
weights are stored as Lagrange-interpolation grids.

The central values of the renormalisation and factorisation scales, $\mu_\mathrm{R}$ and
$\mu_\mathrm{F}$, are chosen in a process-specific way, as discussed in sec.~\ref{subsec:processes_and_measurements}. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted on-the-fly in the Monte Carlo generation upon scale
variations, with the technique presented in Ref.~\cite{Frederix:2011ss}. To this purpose, we use the default \textsc{MG5\_aMC}
implementation, whereby the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed,
see equation~\eqref{eq:}. However, we note that
\textsc{PineAPPL} allows the user to determine the envelope with any point
prescription, see appendix~\ref{app:pineappl-demo} for an example.

@@

The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements,
see secion~\ref{sec:subsec:processes_and_measurements}.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematics observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order of the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of \enquote{standard} EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

@@

When presenting the results, for each of the processes and measurements considered, we compute differential
cross sections for the observables defined in the experimental analyses in two different
ways: directly, by means of \textsc{MG5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and EW couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of \SI{0.1}{\permille} on the integrated cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.

>>>>

In this section we demonstrate the capabilities of \textsc{mg5\_aMC}+\textsc{PineAPPL} by
computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW,
for a representative set of processes in which EW corrections may have a
sizeable effect on the accuracy of the theoretical predictions.
In order to consider some realistic kinematics for these
processes, we rely on a set of measurements that are commonly included in PDF
fits. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We describe
the settings employed for the computations, and the corresponding results for
each process.

@@

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc mg5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, photon-induced (PI) contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation. Monte Carlo
weights are stored as Lagrange-interpolation grids.

The central values of the renormalisation and factorisation scales, $\mu_\mathrm{R}$ and
$\mu_\mathrm{F}$, are chosen in a process-specific way, as discussed in sec.~\ref{subsec:processes_and_measurements}. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted on-the-fly in the Monte Carlo generation upon scale
variations, with the technique presented in Ref.~\cite{Frederix:2011ss}. To this purpose, we use the default \textsc{mg5\_aMC}
implementation, whereby the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed,
see equation~\eqref{eq:9pt}. However, we note that
\textsc{PineAPPL} allows the user to determine the envelope with any point
prescription, see appendix~\ref{app:pineappl-demo} for an example.

@@

The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements,
see section~\ref{subsec:processes_and_measurements}.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematic observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order of the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of \enquote{standard} EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

@@

When presenting the results, for each of the processes and measurements considered, we compute differential
cross sections for the observables defined in the experimental analyses in two different
ways: directly, by means of \textsc{mg5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and EW couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of \SI{0.1}{\permille} on the integrated cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.
